Discussion Draft

### 2020-04-04 doc/Discussion.rmd
cleaning up chromatin str section, add refs

The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons. The COM model (chromosome pairing based model), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the rapid prophase movements (RPMs) (hulten, Rubin et al 2020).

Under **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I. <1> the irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (referred to as along the tetrad arms (cite)).

<2> The number and placement of crossovers alter the distribution of sister cohesion and the 'chromosome structure' which is a consequence of kinetochores attaching to the MT-spindle and tension applied. Single crossover bivalents form what is known as cruciform tetrad in some systems (figure x). This area is reduced when crossover positions are shifted towards the telomere ends of the chromosome. The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.

<3> Conserved feature of gametogenesis -- is the length of timing of the cell cycle (difference between spermat.
faster spermatogenesis -- may select for synchronize the separation homologs or faster separate after alignment at metaphase. While in oogenesis, the longer timing and arrests may select for overall more sister cohesion (a different homolog structures)

<In oocytes the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.>

@@

Our results from MSM and PWD show that the SC length compaction is independent (of the direction of heterochiasmy. Two strains have have more crossovers in males, but females still have longer SC axis. **This indecates that the sexual dimorphism in chromosome organization / DNA compaction is a conserved part of recombination landscape / heterochiasmy**

The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013].

Very few models make predictions for how chromatin organization relates to variation in recombination rates / the recombination landscape.  (the chromatin organization is how chromosomes with the same number of base pairs) are transformed into bottle brush structures with a central axis and most of the genetic material arranged as loops emanating outwards (figure). 
Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume.  This prediction / model might fit broader pattern (such as in At where pollen is the larger cell and has longer axis length) [@cahoonLibuda2019].

>>>>

The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons. The COM model (chromosome pairing based model), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the rapid prophase movements (RPMs) [@hulten2011_COM, @rubin2020mixing].

Under the **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I. The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (cite). The number and placement of crossovers alter the distribution of sister cohesion and the resulting 'chromosome structure' when bivalents are aligned on the metaphase plate (cite). The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.	Differing lengths of the timing of cell cycle between oogenesis and spermatogenesis may impose the the different selective pressures. Faster spermatogenesis may select for synchronization of the separation homologs or faster separatation at the metaphase to anaphase transition. While in oogenesis, the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.

@@

Our results from musculus^MSM^ and musculus^PWD^ show that the SC length and chromatin compaction is uncoupled from the direction of heterochiasmy. These results slightly depart from predictions which nominate chromatin organization (axis length) as the primary driver of heterochiasmy [@petkov2007].  While the relationship may hold for more heterochiasmy patterns, but these results indicate that rapid male specific  evolution in gwRR, is through a different trait than chromatin compaction).

<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>

Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume. This prediction / model might fit broader pattern such as in At where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].

### 2020-04-04 doc/Discussion.tex
cleaning up chromatin str section, add refs

---
title: "Discussion Draft"
date: ''
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

```{r format.Notes, include=FALSE}
#table, wite by all of the columns and rows

#over-view of the synthesis of hypothesis testing
#(this is how we're synthsizing / testing / the framework were using to compare the results to the literature)

#constrast on model types
#-evolution
#-models accounting for chromosome based traits

#HetC traits
# 1. typical landscape
# 2. female variance
# 3. chromatin organization

#faster male evolution traits
# 1. Reversed Direction of Heterochiasmy
# 2. correlated gwRR increase and interference strength

#go through each traits and point out which models support each of the traits
```

```{r Model.testing, include=FALSE}

#imortant of testable hypotheses

#model not included in this comparison

```

We performed hypothesis testing of models from the literature and adapted here fit the main results from this study. We choose to examine three modifier models and two functional and cell physiology models (table X). It is a bit challenging to test all of these models together since they were built with difference parameters and built to describe different patterns of recombination variation. The modifier models were built to explain heterochiasmy variation between sexes and the COM functional model was built to describe the interference pattern and difference in the recombination landscape between the sexes. In cases where the models can't be extended to predicting results, we list as 'NA'.

# Reversed heterochiasmy direction

```{r reversed.HetC, echo=FALSE}

#our results of reversed hetC are novel due to the short evolutionary disances across the panel of mouse strains

#spindle selection modle points

```

@@

The prediction from two of the modifier models, haploid selection and SACE, state that males will generally evolve lower recombination rates. Male biased heterochiasmy is predicted under the two locus modifier model when XXXX.  <does parameterize cases for male biased heterochiasmy, dependent on the stage of action for the driver (MI or MII) and the linkage status (see table 1 Brandvain and Coop).>

There are no prediction for male biased heterochiasmy evolution direction under the COM <pairing model>, but it should be noted that this model was not designed to incorporate this type of data. This model was design to explain the positive interference pattern for of the recombination landscape by incorporating a cell mechanics of rapid prophase movements (RPM) that are thought to facilitate homolog pairing in many species.

Under the spindle based model faster male evolution in male genome wide recombination rates is primarily driven the strength of the spindle assembly checkpoint (SAC) <acting on the chromosome structure>.

The genes and pathways are will known, we refer the reader to [@laneKauppi2019].

@@

Unstable bivalents are detected on the MI spindle, the SAC delays the cell cycle which potentially allows time for lose bivalents to correctly attach to the MT spindle. If the 'unattached' persists apoptosis will be triggered. Genetic variants which increase the number of unstable bivalents, will be selected against in favor for genetic variants causing faster stabilization of bivalents on the MI spindle. Thus the SAC can lead to stabilizing selection on the chromosome structure.

Sperm have lower anueploidy rates compared to oocytes (other lines of evidence) indicating that the 'strength' of the SAC varies between oogenesis and spermatogenesis. In sperm single unpaired bivalents can trigger the SAC. While in females more oocytes progress through meiosis even with unpaired bivalents (refereed to as leaky SAC).

@@

<leaky oocyte SAC linked to acentrosome spindle (multiple aMTOCs) and cell volume, these are features which differ in spermatocyttes (small, centrosome spindle, 2MTOCs)>

Directional selection could be the result in spermatogenesis if the fitness landscape for 'chromosome structure' is altered via genetic architecture changes or GxE effects. (in oogenesis the leaky SAC has weak selection coefficient)

# Greater between cell variation

```{r within.mouse.var, include=FALSE}
#review results, other data

#2. Greater between cell variation in females, leaky SAC in oocytes can cause more between cell variation within an individual.

#2a. relaxed selection on chromosome structure in oocyte / leaky SAC, can lead to larger variance in crossover number between cells. The opposite dynamics of the purifying selection for 'strict' SAC in spermatocytes.

#?2b. (leaky SAC could be adaptive/prevent to selfish genetic elements (Lampson))?

```

Our results of greater between cell variation in females, these results are supported by human and mouse (Gruhn, Lynn Kong). [@lynn2002].

@@

The opposite dynamics of the purifying selection for the 'strict'  SAC in spermatocytes is relaxed selection from a 'permissive' SAC. As mentioned above a fundamental sex difference in the metaphase spindle is the presence of centrioles in spermatocytes which anchors and nucleates the microtubulue spindle at the two cell poles. While oocytes lack centrioles which results in a difuse network of microtubules with mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007). This or larger cell volumne may lead/be a potnetial source of the leaky SAC in oocytes. Leaky SAC, relaxed selection on the anaphase to metaphase transition could increase the between cell variance for crossover number

# Typical landscape

```{r typical.landscape.outline, include=FALSE}

# typical landscape is conserved -- sexually dimorphic meiotic trait (Sardell)

# models

#<spindle model -- the telomeric placement of crossovers, minimizes the amount of sister cohesion connecting homologs>
#3. Sexual dimorphism in broad scale recombination landscape,  telomere bias vs uniform recombination landscapes fit with purifying selection and relaxed selection respectively.
#3a. There are sex differences in demand for chromosome cohesion at MI, the trait / variable causing variation in fitness landscape across chromosome structures might be the amount of sister-cohesion connecting homologs.  In oocytes the cell cycle is slower and arrested in some species, this likely selects for chromosome structures that have stability on the spindle. (required to start with more since the cohesion isn't replenished. The optimal chromosome structure might have more sister cohesion, however given the leaky SAC, selection may not be effective.

#3c. Logic of number and placement modulating this area of sister cohesion and chromosome structure (I plan to make a figure to illustrate)
#-For single crossover acrocentric chromosome the sister cohesion distal to the crossover connects homologs. This area reduced when crossover position shifted to telomere (each end of chromosomes). Two crossover acrocentric chromosome the area is distal from the first until the second crossover.

#3d. Metaphase to anaphase transition proceeds by proteolytic decay of sister cohesion on 'arms'.
#- The telomere bias could have an effect of synchronizing the separation of all the bivalents within a cell.

#*caveats: Type II crossovers not accounted for and chromatid interference effects on (multi-crossover chromsomes) could alter the sister cohesion pattern.

```

Our results on the broad scale pattern of crossover placement are supported by the literature. As reviewed in [@sardell_sex_2020], sexual dimorphism in the broad scale recombination landscape is a highly conserved trait.

The **two locus modifier** and **SACE** model make predictions for sexual dimorphism in the broad scale recombination landscape, for diminishing the effect of drive systems and maintaining larger chromosome blocks respectively.

The two mechanical models also both predict sexual dimorphism in the recombination landscape for distinct reasons. The COM model (chromosome pairing based model), predicts the sex differences in the positioning of crossovers is due to a combination of the length of the axis and intensity of the rapid prophase movements (RPMs) [@hulten2011_COM, @rubin2020mixing].

Under the **spindle based model** we hypothesize that the sexual dimorphism in the recombination landscapes hinges on sex differences on the requirements for chromosome cohesion in late meiosis I. The irreversible process of the metaphase to anaphase transition is initiated by the protyltic decay of the sister cohesion connecting homologs (cite). The number and placement of crossovers alter the distribution of sister cohesion and the resulting 'chromosome structure' when bivalents are aligned on the metaphase plate (cite). The male crossover telomere bias could be adaptive in order to synchronize (or minimize) the timing of homologs separating at metaphase I.	Differing lengths of the timing of cell cycle between oogenesis and spermatogenesis may impose the the different selective pressures. Faster spermatogenesis may select for synchronization of the separation homologs or faster separatation at the metaphase to anaphase transition. While in oogenesis, the slower cell cycle and prolonged arrest may require chromosome structures with greater stability on the MI spindle.

# Chromatin Structure

```{r chromatin.organization, include=FALSE}

#supporting results from lit
#chromatin organization

#models
# none of the models make predictions for the axis length (chromatin organization)
# only COM predicts that sexual dimorphism in axis length is due to cell vol

```

Our results from musculus^MSM^ and musculus^PWD^ show that the SC length and chromatin compaction is uncoupled from the direction of heterochiasmy. These results slightly depart from predictions which nominate chromatin organization (axis length) as the primary driver of heterochiasmy [@petkov2007].  While the relationship may hold for more heterochiasmy patterns, but these results indicate that rapid male specific  evolution in gwRR, is through a different trait than chromatin compaction).

<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>

Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume. This prediction / model might fit broader pattern such as in At where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].

# Evolution of Interference

```{r evo.interference, include=FALSE}
#review results,
#why unexpected, unexpected pattern is great for identifying most parimonious model / explanation
#(we're gonna spend alot of space on this section)

# more convincing that the pattern is real

#lil background, for interferences 
#assuptions for predictions (most modifer models, )

#models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)

#the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)

#**The rapid evolution of the genome wide rate in our high rec strains is PARTLY due to more enrichment of 2CO which have distinct rec landscape from the 2COs in low rec strains.**

#5. Interference evolution shows a positive correlation with gwRR
#5a. Stronger interference for two crossover chromosomes, increases the amount of sister-cohesion connecting the homologs.
#5b. Under spindle selection model, the fitness landscape of optimal chromosome structure may have changed in the high recombining strains resulting in more 2CO, with larger IFD (sister cohesion area).

```

**we find support that interference strength has evolved in the two groups of male strains. There is a small number of positive correlations between genome wide recombination rate and interference strength in the literature.**

@@

Our results support an opposite pattern predicted from the 'logical model', a positive correlation between interference strength and genome wide recombination rates; The within sex comparison of two breeds of cattle with different genome wide recombination rates  [@ma2015_cattle], between lab and wild mice of Peromyscus leucopus, [@peterson2019].

Logically a negative correlation is expected; increasing the number of crossovers across chromosomes would most logically be done by  more densely spacing crossovers along chromosomes and decreasing interference strength. This pattern has empirical support from the most species [@ottoPaysuer2019] and fits well with  the fundamental relationship between the SC area or axis length, the physical upper limit for the number of crossovers. **although it is not particulary strong in our results** We argue that this is due to difference in scale (genome wide vs single chromosome)

@@

Models from Goldstein et al (review in [@ottoPaysuer2019] suggest that if this pattern is widespread interference evolves whenever increased recombination rates evolve. Perhaps a distinguishing feature of models which come to this finding is that the number of crossovers is kept constant. The space across multiple loci (veller?) or between multiple crossovers increases in a positive manner with genome wide recombination rates. Given that the empirical range of crossovers per chromosome is quite small (1-3 [@ottoPaysuer2019, @stapley_variation_2017] and  the obligate crossover rule, the assumption of constraining the number of crossovers per chromosome fits well with empirical data.

@@

# REVIEW

@@

  - It matches fine scale differences in how DSB hotspots had different sex specific strengths with varied due to sex, epigenetc landscape and the gene architecture (Brick et al)

 - incorporates know cellular / molecular mechanisms with a very common pattern sexual dimorphic recombination rate landscape.

- extends modifer model to explain a chromosome based / level pattern
 _weaknesses_, and ways the model diverges from our results

 - the simulation present in the paper are too weak to support the rapid spread of such a modifer system (selection coefficients are too weak to support the evolution -- spread of such a modifer.)
- Other ways this model departs from our results, 
-Always predicting male will be lower recombination rates (or risk breaking up the epistatsis blocks) -- which doesn't fit with our results of faster male specific evolution in gwRR.
However / while not explicitaly arameterized or predicted in this model,
 -- this model predicts larger chuncks of chromosomes will segregate together
 (as outlined in Veller et al)
our results of stronger male interference and stronger interference in high rec strains have the same consequences as predited from this modle

```{r final.review.outline, echo=FALSE}
# comes down to explaning the interference correlation

# spindel selection -- 
#good for sex differences because it frames heterochiasmy within the context of meiosis

#can fit the interference evolution due to the (pressumed) effect on bivalent shape

#caveats and unanswered parts

# SACE is runner up

```

#Future Steps

@@

#blah

@@

#References

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Discussion Draft},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Discussion Draft}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\begin{document}
\maketitle

We performed hypothesis testing of models from the literature and
adapted here fit the main results from this study. We choose to examine
three modifier models and two functional and cell physiology models
(table X). It is a bit challenging to test all of these models together
since they were built with difference parameters and built to describe
different patterns of recombination variation. The modifier models were
built to explain heterochiasmy variation between sexes and the COM
functional model was built to describe the interference pattern and
difference in the recombination landscape between the sexes. In cases
where the models can't be extended to predicting results, we list as
`NA'.

\section{Reversed heterochiasmy
direction}\label{reversed-heterochiasmy-direction}

@@

The prediction from two of the modifier models, haploid selection and
SACE, state that males will generally evolve lower recombination rates.
Male biased heterochiasmy is predicted under the two locus modifier
model when XXXX.

There are no prediction for male biased heterochiasmy evolution
direction under the COM , but it should be noted that this model was not
designed to incorporate this type of data. This model was design to
explain the positive interference pattern for of the recombination
landscape by incorporating a cell mechanics of rapid prophase movements
(RPM) that are thought to facilitate homolog pairing in many species.

Under the spindle based model faster male evolution in male genome wide
recombination rates is primarily driven the strength of the spindle
assembly checkpoint (SAC) .

The genes and pathways are will known, we refer the reader to (Lane and
Kauppi 2019).

@@

Unstable bivalents are detected on the MI spindle, the SAC delays the
cell cycle which potentially allows time for lose bivalents to correctly
attach to the MT spindle. If the `unattached' persists apoptosis will be
triggered. Genetic variants which increase the number of unstable
bivalents, will be selected against in favor for genetic variants
causing faster stabilization of bivalents on the MI spindle. Thus the
SAC can lead to stabilizing selection on the chromosome structure.

Sperm have lower anueploidy rates compared to oocytes (other lines of
evidence) indicating that the `strength' of the SAC varies between
oogenesis and spermatogenesis. In sperm single unpaired bivalents can
trigger the SAC. While in females more oocytes progress through meiosis
even with unpaired bivalents (refereed to as leaky SAC).

@@

Directional selection could be the result in spermatogenesis if the
fitness landscape for `chromosome structure' is altered via genetic
architecture changes or GxE effects. (in oogenesis the leaky SAC has
weak selection coefficient)

\section{Greater between cell
variation}\label{greater-between-cell-variation}

Our results of greater between cell variation in females, these results
are supported by human and mouse (Gruhn, Lynn Kong). (Lynn et al. 2002).

@@

The opposite dynamics of the purifying selection for the `strict' SAC in
spermatocytes is relaxed selection from a `permissive' SAC. As mentioned
above a fundamental sex difference in the metaphase spindle is the
presence of centrioles in spermatocytes which anchors and nucleates the
microtubulue spindle at the two cell poles. While oocytes lack
centrioles which results in a difuse network of microtubules with
mulitple microtubule organizing centers (aMTOCs) (Schuh Ellenburg 2007).
This or larger cell volumne may lead/be a potnetial source of the leaky
SAC in oocytes. Leaky SAC, relaxed selection on the anaphase to
metaphase transition could increase the between cell variance for
crossover number

\section{Typical landscape}\label{typical-landscape}

Our results on the broad scale pattern of crossover placement are
supported by the literature. As reviewed in (Sardell and Kirkpatrick
2020), sexual dimorphism in the broad scale recombination landscape is a
highly conserved trait.

The \textbf{two locus modifier} and \textbf{SACE} model make predictions
for sexual dimorphism in the broad scale recombination landscape, for
diminishing the effect of drive systems and maintaining larger
chromosome blocks respectively.

The two mechanical models also both predict sexual dimorphism in the
recombination landscape for distinct reasons. The COM model (chromosome
pairing based model), predicts the sex differences in the positioning of
crossovers is due to a combination of the length of the axis and
intensity of the rapid prophase movements (RPMs) (Hultén 2011, Rubin,
Macaisne, and Huynh (2020)).

Under the \textbf{spindle based model} we hypothesize that the sexual
dimorphism in the recombination landscapes hinges on sex differences on
the requirements for chromosome cohesion in late meiosis I. The
irreversible process of the metaphase to anaphase transition is
initiated by the protyltic decay of the sister cohesion connecting
homologs (cite). The number and placement of crossovers alter the
distribution of sister cohesion and the resulting `chromosome structure'
when bivalents are aligned on the metaphase plate (cite). The male
crossover telomere bias could be adaptive in order to synchronize (or
minimize) the timing of homologs separating at metaphase I. Differing
lengths of the timing of cell cycle between oogenesis and
spermatogenesis may impose the the different selective pressures. Faster
spermatogenesis may select for synchronization of the separation
homologs or faster separatation at the metaphase to anaphase transition.
While in oogenesis, the slower cell cycle and prolonged arrest may
require chromosome structures with greater stability on the MI spindle.

\section{Chromatin Structure}\label{chromatin-structure}

Our results from musculus\textsuperscript{MSM} and
musculus\textsuperscript{PWD} show that the SC length and chromatin
compaction is uncoupled from the direction of heterochiasmy. These
results slightly depart from predictions which nominate chromatin
organization (axis length) as the primary driver of heterochiasmy
(Petkov et al. 2007). While the relationship may hold for more
heterochiasmy patterns, but these results indicate that rapid male
specific evolution in gwRR, is through a different trait than chromatin
compaction).

Only the COM model predicts sexual dimorphism in chromatin structure
will be longer in oocytes due to greater cell volume. This prediction /
model might fit broader pattern such as in At where pollen is the larger
cell and has longer axis length (Cahoon and Libuda 2019).

\section{Evolution of Interference}\label{evolution-of-interference}

\textbf{we find support that interference strength has evolved in the
two groups of male strains. There is a small number of positive
correlations between genome wide recombination rate and interference
strength in the literature.}

@@

Our results support an opposite pattern predicted from the `logical
model', a positive correlation between interference strength and genome
wide recombination rates; The within sex comparison of two breeds of
cattle with different genome wide recombination rates (Ma et al. 2015),
between lab and wild mice of Peromyscus leucopus, (Peterson, Miller, and
Payseur 2019).

Logically a negative correlation is expected; increasing the number of
crossovers across chromosomes would most logically be done by more
densely spacing crossovers along chromosomes and decreasing interference
strength. This pattern has empirical support from the most species (Otto
and Payseur 2019) and fits well with the fundamental relationship
between the SC area or axis length, the physical upper limit for the
number of crossovers. \textbf{although it is not particulary strong in
our results} We argue that this is due to difference in scale (genome
wide vs single chromosome)

@@

Models from Goldstein et al (review in (Otto and Payseur 2019) suggest
that if this pattern is widespread interference evolves whenever
increased recombination rates evolve. Perhaps a distinguishing feature
of models which come to this finding is that the number of crossovers is
kept constant. The space across multiple loci (veller?) or between
multiple crossovers increases in a positive manner with genome wide
recombination rates. Given that the empirical range of crossovers per
chromosome is quite small (1-3 (Otto and Payseur 2019, Stapley et al.
(2017)) and the obligate crossover rule, the assumption of constraining
the number of crossovers per chromosome fits well with empirical data.

@@

\section{REVIEW}\label{review}

@@

\begin{itemize}
\item
  It matches fine scale differences in how DSB hotspots had different
  sex specific strengths with varied due to sex, epigenetc landscape and
  the gene architecture (Brick et al)
\item
  incorporates know cellular / molecular mechanisms with a very common
  pattern sexual dimorphic recombination rate landscape.
\item
  extends modifer model to explain a chromosome based / level pattern
  \emph{weaknesses}, and ways the model diverges from our results
\item
  the simulation present in the paper are too weak to support the rapid
  spread of such a modifer system (selection coefficients are too weak
  to support the evolution -- spread of such a modifer.)
\item
  Other ways this model departs from our results, -Always predicting
  male will be lower recombination rates (or risk breaking up the
  epistatsis blocks) -- which doesn't fit with our results of faster
  male specific evolution in gwRR. However / while not explicitaly
  arameterized or predicted in this model, -- this model predicts larger
  chuncks of chromosomes will segregate together (as outlined in Veller
  et al) our results of stronger male interference and stronger
  interference in high rec strains have the same consequences as
  predited from this modle
\end{itemize}

\section{Future Steps}\label{future-steps}

@@

\section{blah}\label{blah}

@@

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-cahoonLibuda2019}{}
Cahoon, Cori K, and Diana E Libuda. 2019. ``Leagues of Their Own:
Sexually Dimorphic Features of Meiotic Prophase I.'' \emph{Chromosoma}.
Springer, 1--16.

\hypertarget{ref-hulten2011_COM}{}
Hultén, Maj A. 2011. ``On the Origin of Crossover Interference: A
Chromosome Oscillatory Movement (Com) Model.'' \emph{Molecular
Cytogenetics} 4 (1). Springer: 10.

\hypertarget{ref-laneKauppi2019}{}
Lane, Simon, and Liisa Kauppi. 2019. ``Meiotic Spindle Assembly
Checkpoint and Aneuploidy in Males Versus Females.'' \emph{Cellular and
Molecular Life Sciences} 76 (6). Springer: 1135--50.

\hypertarget{ref-lynn2002}{}
Lynn, Audrey, Kara E Koehler, LuAnn Judis, Ernest R Chan, Jonathan P
Cherry, Stuart Schwartz, Allen Seftel, Patricia A Hunt, and Terry J
Hassold. 2002. ``Covariation of Synaptonemal Complex Length and
Mammalian Meiotic Exchange Rates.'' \emph{Science} 296 (5576). American
Association for the Advancement of Science: 2222--5.

\hypertarget{ref-ma2015_cattle}{}
Ma, Li, Jeffrey R O'Connell, Paul M VanRaden, Botong Shen, Abinash
Padhi, Chuanyu Sun, Derek M Bickhart, et al. 2015. ``Cattle Sex-Specific
Recombination and Genetic Control from a Large Pedigree Analysis.''
\emph{PLoS Genetics} 11 (11). Public Library of Science.

\hypertarget{ref-ottoPaysuer2019}{}
Otto, Sarah P, and Bret A Payseur. 2019. ``Crossover Interference:
Shedding Light on the Evolution of Recombination.'' \emph{Annual Review
of Genetics} 53. Annual Reviews: 19--44.

\hypertarget{ref-peterson2019}{}
Peterson, April L, Nathan D Miller, and Bret A Payseur. 2019.
``Conservation of the Genome-Wide Recombination Rate in White-Footed
Mice.'' \emph{Heredity} 123 (4). Nature Publishing Group: 442--57.

\hypertarget{ref-petkov2007}{}
Petkov, Petko M, Karl W Broman, Jin P Szatkiewicz, and Kenneth Paigen.
2007. ``Crossover Interference Underlies Sex Differences in
Recombination Rates.'' \emph{Trends in Genetics} 23 (11). Elsevier:
539--42.

\hypertarget{ref-rubin2020mixing}{}
Rubin, Thomas, Nicolas Macaisne, and Jean-René Huynh. 2020. ``Mixing and
Matching Chromosomes During Female Meiosis.'' \emph{Cells} 9 (3).
Multidisciplinary Digital Publishing Institute: 696.

\hypertarget{ref-sardell_sex_2020}{}
Sardell, Jason M., and Mark Kirkpatrick. 2020. ``Sex Differences in the
Recombination Landscape.'' \emph{The American Naturalist} 195 (2):
361--79. doi:\href{https://doi.org/10.1086/704943}{10.1086/704943}.

\hypertarget{ref-stapley_variation_2017}{}
Stapley, Jessica, Philine G. D. Feulner, Susan E. Johnston, Anna W.
Santure, and Carole M. Smadja. 2017. ``Variation in Recombination
Frequency and Distribution Across Eukaryotes: Patterns and Processes.''
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences} 372 (1736): 20160455.
doi:\href{https://doi.org/10.1098/rstb.2016.0455}{10.1098/rstb.2016.0455}.

\end{document}
